\bf Enhancing Scheduling Robustness with Task Completion Feedback and Resource Requirement Biasing

### 2019-04-27 paper.tex
Incorporate Dr Antonio's comments in wordsmithing and make all figures have the same y axis range.

\title{\bf Enhancing Scheduling Robustness with Feedback and Biasing}

@@

\begin{abstract}
In previous work the performance of scheduling algorithms for dynamic
scheduling of tasks in a distributed system were evaluated for their
robustness to error in the model of tasks' information.  There it was
found that incorporating task completion timings from the actual distributed
system into the algorithms' model of the system was crucial for achieving
robustness.  In this paper, various degrees of feedback, rather than simply
all-or-none, are evaluated using the same simulated studies as in previous
work and a proposed strategy for biasing model tasks' information is proposed
in order to counteract the most egregious effects of model error on
performance.
\end{abstract}

@@

Such algorithms for dynamic scheduling may be based on heuristics for selecting
which tasks to prioritize execution of and determining when to begin their
execution and on which machine. Other algorithms may actively attempt to
optimize scheduling decisions with respect to a desired outcome based on a
user-defined objective.  Both types of algorithms are generally measured
and compared to one another against such objectives such as minimizing makespan
(time required for completed execution of all tasks) \cite{stochastic}, or, as
in this paper, the degree to which all tasks of a DAG are completed by a
DAG-associated deadline.

Scheduling algorithms may orthogonally be evaluated based on their robustness,
for example, how well the same objective is achieved when information provided
to the scheduling algorithm contains errors such as inaccuracies in the
amount of resource requirements of the tasks.  In previous work \cite{pdpta18}
four dynamic scheduling algorithms' robustness to error with respect to
performance against an objective of completing DAGs before their deadline was
presented, showing how some algorithms were not robust to even the smallest
amount of simulated error.  Additionally, the use of task completion time
feedback from the actual distributed system back into the modeled system used
by the scheduling algorithms was found to substantially improve robustness of
all four scheduling algorithms to even large amounts of simulated error.

The remainder of the paper is organized in the following manner.  Section
\ref{sec:Framework} describes the problem domain and the simulation software's
use of modeling the distributed system where errors in task requirements may
be inherent.  Section \ref{sec:Biasing} presents an approach to counteracting
model error to prevent scheduling algorithms from over-committing system
resources to executing too many tasks concurrently. Section \ref{sec:Results}
presents results of simulated case studies within that software simulator.
Finally, Section \ref{sec:Summary} summarizes the findings from these
simulations and presents the conclusions of our work.

\section{Problem Domain and Simulation Environment}
\label{sec:Framework}

The present work is an extension to previous work \cite{pdpta18} in which a
model-based approach to dynamically scheduling tasks from DAGs (called
workflows) was introduced.  Four scheduling algorithms were tested in a
simulation environment \cite{soasim} given the same 24-hour simulated period
of arriving workflows ranging in size from 5 tasks up to 800.  Each workflow
had a known, predetermined deadline and scheduling algorithms were evaluated
based both on the number of workflows completed before their deadline and the
distribution of workflow counts completed at various normalized proportions
past their deadline (e.g., workflows up to 100\% late relative to the amount
of time between their arrival and deadline).

Scheduling algorithms are used to determine from a given queue of tasks
ready to begin executing (i.e. tasks of arrived workflows that have no
precedence constraints or whose precedent tasks have all completed) both
when the task should begin executing and on which machine of the platform.
Unlike some scheduling research, tasks are permitted to executing concurrently
with other tasks on the same machine which increases machines' overall load
on resources such as CPU and memory and thereby slows the rate of work on each
executing task.  Original work in \cite{pdpta09} details this non-linear
degradation of rate of work (efficiency) due to concurrent task execution.

Figure \ref{fig:platform} illustrates the various components and general
flow of information within the model-based approach to executing and evaluating
scheduling algorithms.  Whereas the modeled tasks' requirements (CPU load,
required number of CPU cycles, and memory load) may contain errors relative to
the true values of those tasks, the model platform may diverge from the
actual platform in terms of which tasks are completed and which are still
executing.

@@

Four scheduling algorithms were studied in this and previous work.  The first
is First-Come, First-Served (FCFS) which prioritizes scheduling tasks from
workflows which arrived earlier over tasks from workflows that arrived later.
The second is Proportional-Least Laxity First which projects a finish time
of a task's workflow based on the rate of completion of its tasks in the past
and then prioritizes tasks from workflows projected to be completed most tardy
proportional to the overall size of the workflow.  The final two are variants
of the Cost-Minimization Scheduling Algorithm (CMSA) \cite{costmin} which
projects a finish time of a task's workflow (as with PLLF) and uses a cost
function to assign as cost to the projected tardiness.  CMSA is used with two
cost functions: a quadratic cost function and sigmoid cost function.

All four studied scheduling algorithms were shown to be sensitive to
even small amounts of error in modeled tasks.  Each scheduling algorithm
exhibited a decreased percent of workflow completed ahead of their deadline
with the smallest amount of error studied.  For three of the scheduling
algorithms the decrease was substantial, but relatively equal, regardless of
the amount of error.  For the fourth scheduling algorithm, the decrease was
less severe overall and was propotionate to the amount of error.  The
fourth algorithm was thus declared to be somewhat robust to small amounts of
error (less than 0.5\%) but ultimately, like all three others, was not
robust for errors of 1\% or greater.

The main result of \cite{pdpta18} showed that incorporating feedback of
task completions from the actual platform to the model platform thereby
preventing the model platform from modeling a task completing before the
actual task completed dramatically increased robustness of all scheduling
algorithms for even errors up to 90\% (the highest amount studied).  Figure
\ref{fig:underestimating} illustrates why the model platform underestimating
the requirements of a task and thereby modeling it as completed ahead of the
actual task can be so detrimental.

@@

In Figure \ref{fig:underestimating} two tasks, $t_1$ and $t_2$, both have
the same amount of CPU cycles, $C$, required for their completion.  In the
top half is the modeled platform's view of time in which task $t_1$ begins
executing first and when it completed, $t_2$ is scheduled to immediately
begin executing.  However, because the modeled requirements of $t_1$ were
an underestimate of the actual $t_1$'s requirements, when the model of
$t_1$ finishes and $t_2$ is scheduled to begin, the actual platform is not
yet finished with $t_1$ and thus must work concurrently on $t_1$ and $t_2$
for some time until the actual $t_1$ task does complete.  This causes both
$t_1$ and $t_2$ to have actual completion times later than the modeled
completion times because of the unintentionally-overallocated resources of
the machine executing the tasks.  By extension, if another task, say $t_3$,
were to be scheduled to begin after $t_2$ finishes this problem could compound
because $t_2$'s actual completion would delayed by concurrent execution with
$t_1$ and further delayed by concurrent execution with $t_3$.

The previous example illustrates why allowing the model to platform to
inform scheduling algorithms when tasks are complete and machines are
idle (or less loaded) in the presense of error which may underestimate tasks'
true requirements can be so detrimental.  With feedback of every task's
completion from the actual platform the modeled completion times of tasks
may be safely ignored.  However, complete feedback of all tasks' completion
may be impractical in a live system, or may simply be cost-prohibitive.
This paper thus seeks to address the question of whether partial feedback
of tasks' completion times may be sufficient to achieve some level of
robustness to model error.  In addition, based specifically on the
knowledge of the underestimating problem illustrated prior, in Section
\ref{sec:Biasing} that follows a specific approach to counteracting
that problem by biasing model task requirements is proposed.  Results of
both partial feedback and biasing are presented in Section \ref{sec:Results}.

@@

As illustrated in Figure \ref{fig:underestimating}, if the model task
requirements are an underestimate of the actual task's requirements then
scheduling algorithms may schedule future tasks to begin executing
unintentionally-concurrent with other tasks, delaying their completion.
However, if the model is known to have error which may underestimate
task requirements, then the model task requirements could simply be
biased by increasing its value in order to reduce (or ideally, eliminate)
the probability that it underestimates the actual task requirements.
In this section three proposed strategies for biasing model task
requirements are proposed.

The simplest form of biasing would be to add a constant value to each
task requirement.  If the maximum magnitude of error for which a model
task requirement may underestimate the actual task requirement is known,
then that value would be the ideal bias constant value because it would
eliminate the model from ever underestimating task requirements but
bound the worst overestimate to the smallest amount possible for such a
constant value biasing strategy.  Practically, the maximum error magnitude
is unlikely to be known, though because it is the most ideal circumstance
for biasing it is included here and in simulated results of Section
\ref{sec:Results}.  The equation for a simple constant bias value, $C$,
addition to each model task requirement, $\hat{X}$, to yield a biased
model task requirement, $\hat{X^b}$, is given in Eq.
\ref{eq:cbias}.

@@

A more practical approach biasing model task requirements is to assume a bound,
not on the magnitude of error, but on the fraction or percentage of error.  It
may be possible, for example, through analysis of past executions of tasks and
workflows, to estimate the maximum percentage of error for each model task
requirement, $\hat{X}$, relative to the actual task requirement, $X$.  In other
words it may be practical to estimate that each model task requirement is
within $P\%$ of the actual task requirement.  Although that means $\hat{X}$
may under- or overestimate $X$ by as much as $P\%$, in order to prevent
underestimating we can adjust $\hat{X}$ as in Eq. \ref{eq:pbias}, hereafter
referred to as the proportionate bias strategy.

@@

Although the proportionate bias strategy can effectively eliminate
underestimated model task requirements with a judiciously chosen $P\%$
which may require less knowledge about the nature of the error than choosing
a proper magnitude for the constant bias strategy value, $C$, this relaxed
requirement comes at a price.  Specifically, where the constant bias
strategy increases the maximum overestimate of actual task requirement, $X$
inherent in $\hat{X}$ by $C$, the proportionate bias strategy `stretches'
the distribution of $\hat{X^b}$ and yields a much larger range of
overestimates.  This is illustrated in Figure \ref{fig:cpbias} where the
distribution of an example model value is adjusted according to the constant
and proportionate bias strategies.  For both adjustments, an ideal value of
$C$ and $P$ are shown, though practically an ideal value wouldn't be known
and have to be estimated itself.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Constant.pdf}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Proportionate.pdf}
	\end{center}
	\caption{Illustration of probabilities for a model task requirement
		with bound error as a percentage, $P$, of the actual task
		requirement, $X$, and the effect of constant and proportionate
		bias strategies transforming the distribution into
		one that never underestimates the actual term, $X$, given an
		ideal value for $C$ and $P$.  Distribution is assumed to be
		zero-mean and triangular.}
	\label{fig:cpbias}
\end{figure}

In order to achieve a less `stretched' distribution for the biased model task
requirement than the proportionate bias strategy while still maintaining the
need only for an estimated maximum percentage, not magnitude, of error, the
third proposed strategy, known as the simple bias strategy, is given in Eq.
\ref{eq:sbias}.  It fails to eliminate the possibility of underestimating
though it does reduce its propability substantially, but also prevents wildly
overestimating task requirements by decreasing the amount of `stretch' in the
biased term's distribution.  Figure \ref{fig:sbias} illustrates the effect of
the simple bias strategy on the distribution of the biased term.

@@

In this paper we presented the case for why complete and reliable feedback
of scheduled task completions from an actual platform to the model used for
scheduling decision-making was found to be necessary to achieve robust
performance of scheduling algorithms in previous work \cite{pdpta18}.
Furthermore we have demonstrated the use of partial feedback as insufficient
at achieving that robustness, but have proposed the use of biasing the model
parameters to prevent or reduce the likelihood of the model underestimating
task requirements thereby modeling tasks as completing earlier than their
actual completion.

Three biasing strategies were proposed, each with different assumptions of
what knowledge is available about the error inherent in the model and
different tradeoffs of balancing the amout of overestimation vs. likelihood
of underestimation of task requirements.  Through simulated case studies we
demonstrated that each biasing strategy is useful for achieving robustness
of scheduling algorithm performance at different amount of model error.

>>>>

\title{\bf Enhancing Scheduling Robustness with Task Completion Feedback and
Resource Requirement Biasing}

@@

\begin{abstract}
Performance and robustness of dynamic scheduling algorithms are evaluated in
the presence of errors in the tasks' resource requirements.  Previous work
found that incorporating task completion events from the actual distributed
system into the algorithms' model of the system was crucial for achieving
robustness.  In the present, various degrees of feedback, rather than simply
all-or-none, are evaluated using the same simulated studies as in previous
work and a proposed strategy for biasing model tasks' information is proposed
in order to counteract the most egregious effects of model error on
performance.
\end{abstract}

@@

Another dimension in the taxonomy of scheduling algorithms deals with whether
tasks are executed `one at a time' on the machine to which they are assigned.
In the present work, multiple tasks may be executed concurrently on a single
machine.

Algorithms for dynamic scheduling may be based on heuristics for selecting
which tasks to prioritize and determining when to begin their execution and on
which machine. Other algorithms attempt to optimize scheduling decisions with
respect to a desired outcome based on a user-defined objective.  Both types of
algorithms are generally measured and compared to one another against such
objectives such as minimizing makespan (time required for completed execution
of all tasks) \cite{stochastic}, or, as in this paper, the degree to which all
tasks of a DAG are completed by a DAG-associated deadline.

Scheduling algorithms may orthogonally be evaluated based on their robustness,
for example, how well the same objective is achieved when information provided
to the scheduling algorithm contains errors such as inaccuracies in resource
requirements of the tasks.  In previous work \cite{pdpta18} four dynamic
scheduling algorithms' robustness to error with respect to performance against
an objective of completing DAGs before their deadline was presented, showing
how some algorithms were not robust to even the smallest amount of error. The
use of task completion event feedback from the actual distributed system back
into the modeled system used by the scheduling algorithms was found to
substantially improve robustness of all four scheduling algorithms even with
large amounts of error.

The remainder of the paper is organized in the following manner.  Section
\ref{sec:Framework} describes the problem domain and the simulation software's
use of modeling the distributed system in which errors in task requirements may
be present.  Section \ref{sec:Biasing} presents a requirements biasing approach
to counteracting model error to prevent scheduling algorithms from
over-committing system resources to executing too many tasks concurrently.
Section \ref{sec:Results} presents results of simulated case studies.
Finally, Section \ref{sec:Summary} summarizes the findings from these
simulations and presents the conclusions of our work.

\section{Problem Domain and \\
Simulation Environment}
\label{sec:Framework}

The present work is an extension to previous work \cite{pdpta18} in which a
model-based approach to dynamically scheduling tasks from DAGs (called
workflows) was introduced.  Four scheduling algorithms were tested in a
simulation environment \cite{soasim} given a 24-hour simulated period
of arriving workflows ranging in size from 5 tasks up to 800.  Each workflow
has a known, predetermined deadline and scheduling algorithms were evaluated
in \cite{pdpta18} based both on the number of workflows completed before their
deadline and the distribution of workflow counts completed at various
normalized proportions past their deadline (e.g., workflows up to 100\% late
relative to the amount of time between their arrival and deadline).

Scheduling algorithms are used to determine \textemdash from a given queue of tasks
ready to begin executing (i.e. tasks of arrived workflows that have no
precedence constraints or whose precedent tasks have all completed)
\textemdash both when the task should begin executing and on which machine of
the platform. Unlike some scheduling research, tasks are permitted to executed
concurrently with other tasks on the same machine, which increases machines'
overall load on resources such as CPU and memory and thereby slows the rate of
work on each executing task.  Original work in \cite{pdpta09} details this
non-linear degradation of rate of work (efficiency) due to concurrent task
execution.

Figure \ref{fig:platform} illustrates the various components and general
flow of information within the model-based approach to executing and evaluating
scheduling algorithms.  The modeled tasks' requirements (CPU load,
required number of CPU cycles, and memory load) may contain errors relative to
the true values of those tasks; these errors then cause the model platform to
diverge from the actual platform in terms of which tasks are completed and
which are still executing.

@@

Four scheduling algorithms were studied in this and previous work. The first
is First-Come, First-Served (FCFS) which prioritizes scheduling tasks from
workflows that arrived earlier over tasks from workflows that arrived later.
The second is Proportional-Least Laxity First which projects a finish time
of a task's overarching workflow based on the rate of completion of its tasks
in the past and then prioritizes tasks from workflows projected to be
completed most tardy proportional to the overall size of the workflow. The
final two algorithms are variants of the Cost-Minimization Scheduling Algorithm
(CMSA) \cite{costmin}, which projects a finish time of a task's workflow (as
with PLLF) and uses a cost function to assign a cost to the projected
tardiness. CMSA is used with two cost functions: a quadratic cost function and
sigmoid cost function.

All four studied scheduling algorithms were shown to be sensitive to
even small amounts of error in modeled tasks \cite{pdpta18}.  Each scheduling
algorithm exhibited a decreased percent of workflow completed ahead of their
deadline with the smallest amount of error studied.  For three of the
scheduling algorithms the decrease was substantial, but relatively equal,
regardless of the amount of error.  For the fourth scheduling algorithm, the
decrease was less severe overall and was propotionate to the amount of error.
The fourth algorithm was thus declared to be somewhat robust to small amounts
of error (less than 0.5\%) but ultimately, like all three others, was not
robust for errors of 1\% or greater.

The main result of \cite{pdpta18} showed that incorporating feedback of
task completions from the actual platform to the model platform (thereby
preventing the model platform from modeling a task completing before the
actual task completed) dramatically increased robustness of all scheduling
algorithms, even for errors up to 90\% (the highest amount studied) in the
modeled tasks' requirement for amount of work in CPU cycles.  Figure
\ref{fig:underestimating} illustrates why the model platform's underestimation
of the requirements of a task (and thereby modeling it as completed ahead of
the actual task) can be so detrimental.

@@

In Figure \ref{fig:underestimating} two tasks, $t_1$ and $t_2$, both have
the same amount of CPU cycles, $C$, required for their completion.  In the
top half is the modeled platform's view of time in which task $t_1$ begins
executing first and when it completed, $t_2$ is scheduled to immediately
begin executing.  However, because the modeled requirements of $t_1$ were
an underestimate of the actual $t_1$'s requirements, when the model of
$t_1$ finishes and $t_2$ is scheduled to begin, the actual platform is not
yet finished with $t_1$ and thus must work concurrently on $t_1$ and $t_2$
for some time until the actual $t_1$ task does complete.  This causes both
$t_1$ and $t_2$ to have actual completion times later than the modeled
completion times because of the unintentional over allocation of resources of
the machine executing the tasks.  By extension, if another task, say $t_3$,
were to be scheduled to begin after $t_2$ finishes this problem could compound
because $t_2$'s actual completion would be further delayed by the additional
concurrent execution with $t_3$.

The previous example illustrates why allowing the model platform to
inform scheduling algorithms when tasks are complete and machines are
idle (or less loaded) in the presense of error (which may underestimate tasks'
true requirements) can be so detrimental.  With feedback of every task's
completion from the actual platform, the modeled completion times of tasks
may be safely ignored.  However, complete feedback of all tasks' completion
may be impractical in a live system, or may simply be cost-prohibitive.
This paper thus seeks to address the question of whether partial feedback
of tasks' completion times may be sufficient to achieve some level of
robustness to model error.  In addition, based specifically on the knowledge
of the underestimating problem illustrated prior, Section \ref{sec:Biasing}
proposes a specific approach to counteracting that problem by biasing model
task requirements.  Results of both partial feedback and biasing are presented
in Section \ref{sec:Results}.

@@

As illustrated in Figure \ref{fig:underestimating}, if the model task
requirements are an underestimate of the actual task's requirements then
scheduling algorithms may schedule future tasks to begin executing
unintentionally-concurrent with other tasks, delaying their completion.
However, if the model is known to have error which may underestimate task
requirements, then the model task requirements could simply be biased by
increasing its assumed (provided) value in order to reduce (or ideally,
eliminate) the probability that it underestimates the actual task requirements.
In this section three proposed strategies for biasing model task requirements
are proposed.

The simplest form of biasing is to add a constant value to each task
requirement.  If the maximum magnitude of error for which a model
task requirement may underestimate the actual task requirement is known,
then that value would be the ideal bias constant value because it would
eliminate the model from ever underestimating task requirements while
minimizing the amount of overestimation. Practically, the maximum error
magnitude is unlikely to be known. However, because it is the most ideal
circumstance for biasing it is included here and in simulated results of
Section \ref{sec:Results}. The equation for a simple constant bias value, $C$,
addition to each model task requirement, $\hat{X}$, to yield a biased model
task requirement, $\hat{X^b}$, is given in Eq. \ref{eq:cbias}.

@@

A more realistic approach to biasing model task requirements is to assume a
bound, not on the magnitude of error, but on the fraction or percentage of
error. It may be possible, for example, through analysis of past executions of
tasks and workflows, to estimate the maximum percentage of error for each model
task requirement, $\hat{X}$, relative to the actual task requirement, $X$. In
other words it may be practical to estimate that each model task requirement is
within $P\%$ of the actual task requirement.  Although that implies $\hat{X}$
may under- or overestimate $X$ by as much as $P\%$, in order to prevent
underestimating we can adjust $\hat{X}$ as in Eq. \ref{eq:pbias}, hereafter
referred to as the proportionate bias strategy.

@@

Although the proportionate bias strategy can effectively eliminate
underestimated model task requirements with a judiciously chosen $P\%$
which may require less knowledge about the nature of the error than choosing
a proper magnitude for the constant bias strategy value, $C$, this relaxed
requirement comes at a price. Specifically, where the constant bias
strategy shifts the actual task requirement by $C$, the proportionate bias
strategy `stretches' the distribution of $\hat{X^b}$ and yields a much larger
range of overestimates. This is illustrated in Figure \ref{fig:cpbias} where
the distribution of an example model value is adjusted according to the
constant and proportionate bias strategies.  For both adjustments, an ideal
value of $C$ and $P$ are shown, though practically an ideal value wouldn't be
known and have to be estimated itself.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Constant.pdf}
		\includegraphics[width=0.5\textwidth]{figures/BiasVisualization_Proportionate.pdf}
	\end{center}
	\caption{Illustration of probabilities for a model task requirement
		with bound error as a percentage, $P$, of the actual task
		requirement, $X$, and the effect of constant and proportionate
		bias strategies transforming the distribution into
		one that never underestimates the actual term, $X$, given an
		ideal value for $C$ and using the exact value of $P$ (the bounded
		error) in the proportionate bias strategy's Eq. \ref{eq:pbias}.
		Distribution is assumed to be zero-mean and triangular.}
	\label{fig:cpbias}
\end{figure}

In order to achieve a less `stretched' distribution for the biased model task
requirement than the proportionate bias strategy while still maintaining the
need only for an estimated maximum percentage, not magnitude, of error, the
third proposed strategy, known as the simple bias strategy, is given in Eq.
\ref{eq:sbias}. This third strategy fails to eliminate the possibility of
underestimating though it does reduce its propability substantially, but also
prevents wildly overestimating task requirements by decreasing the amount of
`stretch' in the biased term's distribution.  Figure \ref{fig:sbias}
illustrates the effect of the simple bias strategy on the distribution of the
biased term.

@@

The case for why complete and reliable feedback of scheduled task completions
from an actual platform to the model used for scheduling decision-making was
found to be necessary to achieve robust performance of scheduling algorithms,
as first described in previous work \cite{pdpta18}. Furthermore we have
demonstrated the use of partial feedback to be insufficient at achieving
similar robustness, but have proposed the use of biasing the model parameters
to prevent or reduce the likelihood of the model underestimating task
requirements thereby modeling tasks as completing earlier than their actual
completion. Three biasing strategies were proposed, each with different
assumptions of what knowledge is available about the error inherent in the
model and different tradeoffs of balancing the amout of overestimation vs.
likelihood of underestimation of task requirements.  Through simulated case
studies we demonstrated that each biasing strategy is useful for achieving
robustness of scheduling algorithm performance at different amount of model
error.
